---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(iqtigfunctions)
```
# iqtigfunctions

<!-- badges: start -->
<!-- badges: end -->

The goal of `iqtigfunctions` is to provide an implementation of IQTIG specific R functions as used in the published indicator specifications (QIDB). For a detailed, technical reference please use the `help` function, or take a look at the IQTIG R readme (in German), distributed with the QIDB.

## Operators

### Comparison

A number of binary comparison operators have been defined to compare two numbers with a special handling of NAs.

* `%!=%`
* `%<%`
* `%<=%`
* `%==%`
* `%>%`
* `%>=%`

```{r}
5 %<=% 10
2 %==% NA_real_
```

### Between

In addition `%between%` checks element-wise if a vector is within an interval.

```{r}
1:5 %between% c(2, 4)
```

### Matching

Some computional rules check values against a list of codes (e.g. ICD or OPS codes). The following four operators have been defined to make it easier to match values against a list of values. Theses values can also contain so-called wildcards (`%`).

* `%all_in%`
* `%all_like%`
* `%any_in%`
* `%any_like%`

```{r}
data <- data.frame(
  OPSCHLUESSEL_1 = c("5-123.5", "5-134", "5-156"),
  OPSCHLUESSEL_2 = c("5-001", "5-100", NA_character_),
  OPSCHLUESSEL_3 = NA_character_,
  TDS_B = c(1, 2, 3),
  stringsAsFactors = FALSE
)

# the following function gathers all fields spread over multiple columns
# and converts it to a list column
data <- gather_composite_fields(data)
data$OPSCHLUESSEL
```

```{r}
with(data, OPSCHLUESSEL %any_in% c("5-12", "5-123.2", "5-156"))
```

```{r}
with(data, OPSCHLUESSEL %all_like% c("5-123%", "5-1%", "5-168.12:R"))
```

### Grouping

Often it is necessary to compute a value on subsets of records defined by a certain property. For example, the latest date of operation among all operations of a patient.

* `%group_by%`

```{r}
with(mtcars, maximum(hp) %group_by% cyl)
```

```{r}
with(mtcars, maximum(hp) %group_by% c(cyl, drat))
```

## General functions

There is also a small number of general functions used in computational rules.

* `maximum`
* `minimum`
* `replace_na`
* `row_sums`
* `to_year`

## License

MIT
